---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Jiaxuan Song
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false


library(tidyverse)
library(palmerpenguins)
library(rstanarm)
library(modelsummary)
library(janitor)
library(arrow)
library(ggplot2)
library(knitr)
# Load the dataset
data <- read_parquet(here::here("data/02-analysis_data/cchs_cleaned_data.parquet"))
```


# Introduction

Depression is a global mental health challenge that significantly affects the quality of life and economic productivity. According to the World Health Organization (WHO), over 280 million people worldwide experience depression, making it a leading cause of mental disability [@who_depression]. The multifactorial nature of depression—spanning biological, psychological, social, and environmental dimensions—makes it complex to study and address effectively. [@stringaris2017depression] Understanding the key factors influencing depression is critical for developing targeted interventions and policies to mitigate its burden. Depression is understood to result from a combination of factors across biological, psychological, social, and environmental dimensions [@epine2011increasing]. For instance, genetic studies indicate that certain genes may increase an individual’s susceptibility to depression, while sociological research highlights the importance of environmental factors, such as poverty, unemployment, or a lack of social support [@addis1995people]. However, most existing studies focus on isolated variables, overlooking the intricate interplay between these factors. Furthermore, much of the current research is limited to Western contexts, leaving significant gaps in understanding depression across different cultural settings [@chentsova2009understanding].

Although extensive research has been conducted on depression, most studies focus on individual factors such as socioeconomic status, genetics, or social support, without fully exploring the interactions between these variables. Furthermore, much of the existing literature is region-specific, leaving gaps in understanding cross-cultural or generalized patterns. In the Canadian context, while research on depression has increased, there is still a lack of systematic analysis of specific social and economic factors contributing to depression. This project aims to address this gap by analyzing data from the Canadian Community Health Survey (CCHS) for the years 2019 and 2020. [@statcan_cchs] The dataset includes various potential predictors of depression, such as age, gender, income, employment status, social support, and health behaviors. By applying statistical techniques, this study seeks to identify the most critical factors influencing depression risk and their relative contributions, with a specific focus on trends and disparities evident during this period in Canada.

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....






# Data {#sec-data}

## Overview

The dataset used in this study was sourced from the public use microdata file (PUMF) of the Canadian Community Health Survey (CCHS), which provides comprehensive data for health regions and combinations of health regions across Canada. Covering a two-year period, the dataset includes interviews with approximately 130,000 respondents aged 12 or older residing in households across all provinces and territories [@statcan_cchs].

The dataset contains information on a wide range of topics, including physical activity, height and weight, smoking, exposure to second-hand smoke, alcohol consumption, general health, chronic health conditions, injuries, and use of healthcare services. Additionally, it offers insights into socio-demographic, income, and labour force characteristics of the population. For this study, the dataset underwent preprocessing to focus specifically on variables relevant to depression analysis, such as socio-demographic factors, health behaviors, and chronic health conditions. These records were retained to form the basis for the analysis, enabling a detailed examination of factors influencing depression in Canada.

In this project, we employ `R` [@citeR] as the primary statistical programming language, integrating several key packages to enhance analysis and visualization. These include the `tidyverse` [@tidyverse] suite for data manipulation and cleaning, `palmerpenguins` [@palmerpenguins] for illustrative data examples, and `knitr` [@knitr2014] for dynamic report generation. Additionally, we use `rstanarm` [@rstanarm] for linear regression modeling, `modelsummary` [@modelsummary] for streamlined model summaries, `janitor` [@janitor] for data cleaning utilities, `arrow` [@arrow] for efficient data access and processing, and `ggplot2` [@ggplot2] for advanced data visualization.

## Measurement
	
In mt analysis of depression and its associated factors in Canada, I translate complex health, behavioral, and socio-demographic data into a structured dataset. The dataset, derived from the 2019–2020 Canadian Community Health Survey (CCHS), provides a nationally representative sample of respondents aged 12 and older, enabling us to analyze depression severity and its potential predictors. While variables such as depression severity score (DEPDVPHQ) and classification of depression levels (DEPDVSEV) simplify the nuanced experiences of respondents, they serve as practical outcome measures, allowing us to examine patterns of mental health across diverse populations.

To ensure consistency and accuracy, we implemented a systematic data-cleaning process, including handling missing values, removing duplicates, and verifying variable formats. For example, we standardized key variables such as marital status (DHHGMS), categorizing respondents as either "Married/Common-law" or "Other," and grouped age (DHHGAGE) into meaningful ranges (e.g., "12-17," "18-34"). Weekly alcohol consumption (ALWDVWKY) was retained as a measure of behavioral health, while household income (INCDGHH) and education level (EHG2DVH3) provided socio-economic context. By focusing on these relevant variables, we streamlined the dataset to emphasize the factors most likely to influence depression outcomes.

To specifically examine depression, we utilized two key variables: the depression severity score (DEPDVPHQ) and the classification of depression levels (DEPDVSEV), enabling a detailed analysis of both continuous and categorical measures of depression. Data standardization efforts included recoding missing or ambiguous information to ensure clarity. This systematic preparation provides a robust foundation for exploring relationships between predictors and depression severity, allowing us to identify key factors influencing mental health outcomes in Canada and develop actionable insights for intervention and policy.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.



Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false




```

Talk way more about it. 

## Predictor variables
```{r}
library(ggplot2)
library(dplyr)

# Assuming the dataset is named 'data'
# Select and clean relevant columns
cleaned_data <- data %>%
  select(DHHGAGE, DEPDVPHQ, DHHGMS) %>%
  drop_na() %>%
  rename(
    `Age Group` = DHHGAGE,
    `Depression Score` = DEPDVPHQ,
    `Marital Status` = DHHGMS
  ) %>%
  mutate(
    `Marital Status` = case_when(
      `Marital Status` == 1 ~ "Married/Common-law",
      `Marital Status` == 2 ~ "Other",
      TRUE ~ "Unknown"
    ),
    `Age Group` = case_when(
      `Age Group` == 1 ~ "12-17",
      `Age Group` == 2 ~ "18-34",
      `Age Group` == 3 ~ "35-49",
      `Age Group` == 4 ~ "50-64",
      `Age Group` == 5 ~ "65+",
      TRUE ~ as.character(`Age Group`)
    )
  )

# Create the grouped bar chart
ggplot(cleaned_data, aes(x = `Age Group`, y = `Depression Score`, fill = `Marital Status`)) +
  stat_summary(fun = mean, geom = "bar", position = "dodge") +
  labs(
    title = "Depression Score by Age Group and Marital Status",
    x = "Age Group",
    y = "Average Depression Score",
    fill = "Marital Status"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

```{r}
library(ggplot2)
library(dplyr)

# Assuming the dataset is named 'data'
# Clean and process the data
cleaned_data <- data %>%
  select(DEPDVPHQ, INCDGHH, EHG2DVH3) %>%
  drop_na() %>%
  rename(
    `Depression Score` = DEPDVPHQ,
    `Income Group` = INCDGHH,
    `Education Level` = EHG2DVH3
  ) %>%
  mutate(
    `Income Group` = case_when(
      `Income Group` == 1 ~ "Lowest Income",
      `Income Group` == 2 ~ "Low-Middle Income",
      `Income Group` == 3 ~ "Middle Income",
      `Income Group` == 4 ~ "Upper-Middle Income",
      `Income Group` == 5 ~ "Highest Income",
      TRUE ~ "Unknown"
    ),
    `Education Level` = case_when(
      `Education Level` == 1 ~ "No Formal Education",
      `Education Level` == 2 ~ "Some High School",
      `Education Level` == 3 ~ "High School Graduate",
      `Education Level` == 4 ~ "Some College",
      `Education Level` == 5 ~ "College Graduate",
      `Education Level` == 6 ~ "Postgraduate",
      TRUE ~ "Unknown"
    )
  ) %>%
  group_by(`Income Group`, `Education Level`) %>%
  summarise(Average_Depression_Score = mean(`Depression Score`, na.rm = TRUE)) %>%
  arrange(match(`Income Group`, c(
    "Lowest Income", "Low-Middle Income", "Middle Income", 
    "Upper-Middle Income", "Highest Income"
  )))

# Create the line graph with Education Level as a color factor
ggplot(cleaned_data, aes(x = `Income Group`, y = Average_Depression_Score, group = `Education Level`, color = `Education Level`)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  labs(
    title = "Depression Trends Across Income Levels and Education",
    x = "Income Group",
    y = "Average Depression Score",
    color = "Education Level"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))




```

```{r}
library(ggplot2)
library(dplyr)

# Assuming the dataset is named 'data'
# Select and clean relevant columns for alcohol consumption and depression
cleaned_data <- data %>%
  select(ALCDVTTM, DEPDVPHQ) %>%
  drop_na() %>%
  rename(
    `Alcohol Consumption` = ALCDVTTM,
    `Depression Score` = DEPDVPHQ
  ) %>%
  mutate(
    `Alcohol Consumption` = case_when(
      `Alcohol Consumption` == 1 ~ "None",
      `Alcohol Consumption` == 2 ~ "Low",
      `Alcohol Consumption` == 3 ~ "Medium",
      `Alcohol Consumption` == 4 ~ "High",
      TRUE ~ as.character(`Alcohol Consumption`)
    ),
    `Depression Category` = case_when(
      `Depression Score` <= 5 ~ "Low Depression",
      `Depression Score` > 5 & `Depression Score` <= 10 ~ "Moderate Depression",
      `Depression Score` > 10 ~ "Severe Depression",
      TRUE ~ as.character(`Depression Score`)
    )
  )

# Summarize the data to count occurrences of each alcohol consumption and depression category combination
alcohol_depression_summary <- cleaned_data %>%
  count(`Alcohol Consumption`, `Depression Category`) %>%
  mutate(percentage = n / sum(n) * 100)

# Create the bar chart comparing alcohol consumption and depression categories
ggplot(alcohol_depression_summary, aes(x = `Alcohol Consumption`, y = n, fill = `Depression Category`)) +
  geom_bar(stat = "identity", position = "stack") +  # Stacked bar chart
  labs(
    title = "Comparison of Alcohol Consumption and Depression Categories",
    x = "Alcohol Consumption",
    y = "Count",
    fill = "Depression Category"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))




```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2


```



\newpage


# References


